### 2014-02-28 paper/concrete.tex
checkpoint.

The operational semantics we have defined in the previous section
satisfy termination sensitive non-interference by design.
We achieve this general statement that works for a large class of
languages by having different task execute completely isolated from
each other, such that every task has it's own state.
In some cases, this strong separation is desirable, or even necessary.
Languages like C provide direct access to the addresses without
mechanisms in the language to achieve a separation of the heap.
However, for other languages this isolation of tasks can be
undesirable, e.g., for performance reasons.
For instance, for the mini-ML language, our presentation so far
requires a separate heap per task, which is not very practical.
Instead, we would like to
more tightly couple the integration of the target and IFC
language by reusing existing infrastructure.  In the running example,
a concrete implementation might use a single global heap, and ensure
that references do not cross task boundaries, and thus preserving
non-interference.

@@

Instead of using a configuration of the form
\[|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|\]
as is done in |specLangML alpha|
where every task has its
own ML heap, we would like a single global heap as in
|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|.
To keep up a fictional separation between the portions
of the global heap that tasks can use, we forbid references to
cross task boundaries.  To achieve such a system, we can
adapt most of the operational rules easily to work with these
new configurations.  In the abstract system, the \textsc{I-fork}
rule created a fresh heap for the new task to execute in.  Since we
want to forbid references to cross task boundaries, we do not modify
the global heap and instead add an additional premise that the forked
expression |ie| must not contain any references.
Finally, we add the same condition to the \textsc{I-send} rule.
Tasks that try to fork or send an expression that contains a
reference will now get stuck, and will stop executing by the
\textsc{I-noStep} rule.
We call this family of languages |concreteLangMl alpha|, which
is still parametrized by the scheduling policy |alpha|.

>>>>

The operational semantics we have defined in the previous section
satisfy non-interference by design.
We achieve this general statement that works for a large class of
languages by having different task execute completely isolated from
each other, such that every task has it's own state.
In some cases, this strong separation is desirable, or even necessary.
Languages like C provide direct access to the addresses without
mechanisms in the language to achieve a separation of the heap.
However, for other languages this isolation of tasks can be
undesirable, e.g., for performance reasons.
For instance, for the mini-ML language, our presentation so far
requires a separate heap per task, which is not very practical.
Instead, we would like to
more tightly couple the integration of the target and IFC
language by reusing existing infrastructure.  In the running example,
a concrete implementation might use a single global heap.
More precisely, instead of using a configuration of the form
\[|iconf iS (fullconf id1 il1 tS1 ie1, fullconf id2 il2 tS2 ie2 ldots)|\]
we would like a single global heap as in
|oneheapiconf iS tS (oneheapfullconf id1 il1 ie1, oneheapfullconf id2 il2 ie2, ldots)|.

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated.  In particular, it is not
sound to share references to a heap cell between different tasks
in general.  Otherwise, information can be trivially leaked by copying
secret data into a reference in one task, and then
read again by a another task with a lower label.
Clearly, we cannot
modify the semantics in arbitrary ways if we are hoping to preserve
non-interference, but how can we know if our changes to the abstract
semantics preserve non-interference?
The intention of our single heap implementation was to be more efficient
by using one global heap, but conceptually maintain isolation between
task by not allowing sharing of references between tasks.
This intuition of having a different (potentially more efficient)
concrete semantics that behave like the abstract semantics
can be formalized by the following definition:

\begin{definition}[Isomorphism of information-flow control languages]
  A language \con{|(C, .->, erasef l)|} is \textit{isomorphic} to a
  language \abs{|(iC, .->, erasef l)|} if there exist total functions |f
  : tC -> iC| and |finv : iC -> tC| such that |f .. finv = idf tC| |finv
  .. f = idf iC|.  Furthermore, |f| and |finv| are functorial (e.g. if
  $\abs{x\ R\ y}$ then $f(\abs{x})\ \con{R}\ f(\abs{y})$) over both
  $l$-equivalences and |.->|.

  If we weaken this restriction such that |finv| does
  not have to be functorial over |.->|, we call the
  language \con{|(C, .->, erasef l)|} \textit{weakly isomorphic} to
  \abs{|(iC, .->, erasef l)|}.
\end{definition}

Providing an isomorphism between the two languages allows us to
preserve (termination sensitive or insensitive) non-interference
as the following two theorems state.

\begin{theorem}[Weak isomorphism preserves termination insensitive non-interference]
  If a language \con{L} is weakly isomorphic to a language \abs{L}, and \abs{L}
  satisfies TINI, then \con{L} satisfies TINI.
\end{theorem}

A na\"ive attempt at our single heap semantics
will not allow us to construct an isomorphism
with the abstract language |specLangML alpha|, and instead highlights
where things will go wrong.

@@

To keep up a fictional separation between the portions
of the global heap that tasks can use, we forbid references to
cross task boundaries.  To achieve such a system, we can
adapt most of the operational rules easily to work with these
new configurations.  In the abstract system, the \textsc{I-fork}
rule created a fresh heap for the new task to execute in.  Since we
want to forbid references to cross task boundaries, we do not modify
the global heap and instead add an additional premise that the forked
expression |ie| must not contain any references.
Finally, we add the same condition to the \textsc{I-send} rule.
Tasks that try to fork or send an expression that contains a
reference will now get stuck, and will stop executing by the
\textsc{I-noStep} rule.
We call this family of languages |concreteLangMl alpha|, which
is still parametrized by the scheduling policy |alpha|.
